Quantitative Evaluation of Touch-Screen Performance with Transition Matrix Measures

### 2015-09-01 flux-entropy-paper/flux-entropy-paper.tex
cleanup and new faceted plots

\section{A Corpus of Musical Touch-Screen Interactions}

@@

The distribution of flux and entropy with respect to session contexts,
types and instrumentation are shown in Figures \ref{fig:flux-boxplot},
\ref{fig:entropy-boxplot}, and \ref{fig:flux-entropy-distribution}. A
three-way ANOVA procedure was performed on the data set to find
whether the metatdata could significantly predict the entropy and flux
values. In the following analysis, only the 95 performative sessions
were considered (i.e. not the demonstration or data collection
contexts).

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/flux-boxplot}
  \caption{Distributions of flux values by performance context. Study
    and rehearsal sessions had the highest flux for improvisations
    which fell during performances. Compositions had lower flux in
    rehearsals than performances.
    \label{fig:flux-boxplot}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/entropy-boxplot}
  \caption{Distribution of entropy by
  performance context. Demonstration sessions had much lower entropy
  indicating data unevenly spread in the transition matrices. This
  could be because untrained
  participants did not explore the full gestural potential of the
  instruments. 
  \label{fig:entropy-boxplot} }
\end{figure}

@@

Figures \ref{fig:type-section-flux} and \ref{fig:type-section-entropy}
show that the flux and entropy of compositions are much more stable
with respect to section than for improvisations. The stability of
compositions is to be expected since performers are expected to use
the same gestures in each rehearsal and performance of a composition,
it may be that different kinds of compositions could result in
different structural differences. Both measures have a downward trend
through sections in improvisations with Kruskal-Wallis tests showing a
significant effect of section on flux ($\chi^2(2)=5.8, p=0.05$). This
could be explained by performers entering an exploration stage at the
start of an improvisation, where they change gesture frequently to
experiment with different ideas. At the end of an improvisation,
performers may change gesture less frequently while ``winding up'' the
performance.

The gestural change in improvisation can be further explored by
dividing the sessions by performance context. Figures
\ref{fig:context-section-flux}, and \ref{fig:context-section-entropy} show
this division for improvisations only. In rehearsals and studies the
distribution of flux and entropy over sections has a similar shape,
both measures drop in the ending section of studies and after the
beginning section in rehearsals. In performances however, flux drops
after the beginning while entropy drops substantially in the ending
section. This indicates a more constrained performance style in this
closing section, where performers transition between fewer gestures in
the final section of performances.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/type-section-flux}
  \caption{The flux of compositions and improvisations divided into
    beginning, middle, and end. Compositions remain stable throughout,
    but improvisations show a downward trend.
    \label{fig:type-section-flux}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/type-section-entropy}
  \caption{The entropy of compositions and improvisations divided by
    beginning, middle, and end. Composition entropy tends to move up
    at the end, while flux moves down.
    \label{fig:type-section-entropy}}
\end{figure}

% todo add the results from the plots.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/context-section-flux}
  \caption{The distribution in performances divided by ``beginnings,
    middles, and ends''.
    \label{fig:context-section-flux}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/context-section-entropy}
  \caption{The distribution of entropy in performances divided into
    sections and by performance context.
    \label{fig:context-section-entropy}}
\end{figure}

@@

While the interaction session in our corpus have many different
contexts, they are all part of a program of artistic and HCI research
that has evolved over time. Figures \ref{fig:flux-through-time} and
\ref{fig:entropy-through-time} show the values of flux and entropy for
each session in the corpus. Over the course of this research program,
different ensembles and different activities have been emphasised,
some of these differences can be seen as clusters of performances with
similar flux and entropy in the two plots. For instance, the earliest
series of rehearsals and performances had flux between 0.05 and 0.15,
while the period of activity after July 2014 had flux between 0.1 and
0.25. This change corresponded with an intensive series of study
sessions and rehearsals where the participants developed a more
fluctuating style of performance. The entropy between these two
periods, however, remains at a similiar level, showing the the space
of available gestures that the performers explored was more constant.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/flux-through-time}
  \caption{Distribution of flux values of sessions through time.
    Different performance styles are visible as clusters of similar
    flux regardless of performance context.
    \label{fig:flux-through-time}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/entropy-through-time}
  \caption{Distribution of entropy values of sessions through time.
    \label{fig:entropy-through-time}}
\end{figure}

% todo add anymore results from these plots

>>>>

\section{A Corpus of Touch-Screen Interactions}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/context-flux-entropy-boxplot}
  \caption{Distributions of flux and entropy values by performance context. Study
    and rehearsal sessions had the highest flux for improvisations
    which fell during performances. Compositions had lower flux in
    rehearsals than performances. Demonstration sessions had much lower entropy
  indicating data unevenly spread in the transition matrices. This
  could be because untrained
  participants did not explore the full gestural potential of the
  instruments. 
  \label{fig:flux-entropy-boxplot}}
\end{figure}

The distribution of flux and entropy with respect to session contexts,
types and instrumentation are shown in Figures
\ref{fig:flux-entropy-boxplot} and
\ref{fig:flux-entropy-distribution}. A three-way ANOVA procedure was
performed on the data set to find whether the metatdata could
significantly predict the entropy and flux values. In the following
analysis, only the 95 performative sessions were considered (i.e. not
the demonstration or data collection contexts).

@@

@@

Figures \ref{fig:type-section-flux-entropy} shows that the flux and
entropy of compositions are much more stable with respect to section
than for improvisations. The stability of compositions is to be
expected since performers are expected to use the same gestures in
each rehearsal and performance of a composition, it may be that
different kinds of compositions could result in different structural
differences. Both measures have a downward trend through sections in
improvisations with Kruskal-Wallis tests showing a significant effect
of section on flux ($\chi^2(2)=5.8, p=0.05$). This could be explained
by performers entering an exploration stage at the start of an
improvisation, where they change gesture frequently to experiment with
different ideas. At the end of an improvisation, performers may change
gesture less frequently while ``winding up'' the performance.

The gestural change in improvisation can be further explored by
dividing the sessions by performance context. Figure
\ref{fig:context-section-flux-entropy} shows this division for
improvisations only. In rehearsals and studies the distribution of
flux and entropy over sections has a similar shape, both measures drop
in the ending section of studies and after the beginning section in
rehearsals. In performances however, flux drops after the beginning
while entropy drops substantially in the ending section. This
indicates a more constrained performance style in this closing
section, where performers transition between fewer gestures in the
final section of performances.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/type-section-flux}
\includegraphics[width=\linewidth]{figures/type-section-entropy}
  \caption{The flux of compositions and improvisations divided into
    beginning, middle, and end. Composition flux remains stable throughout,
    but improvisations show a downward trend and composition entropy tends to move up
    at the end, while flux moves down.
    \label{fig:type-section-flux-entropy}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/context-section-flux}
  \includegraphics[width=\linewidth]{figures/context-section-entropy}
  \caption{The distribution of flux and entropy in performances divided by ``beginnings,
    middles, and ends''.
    \label{fig:context-section-flux-entropy}}
\end{figure}

@@

While the interaction session in our corpus have many different
contexts, they are all part of a program of artistic and HCI research
that has evolved over time. Figure \ref{fig:flux-entropy-through-time}
shows the values of flux and entropy for each session in the corpus.
Over the course of this research program, different ensembles and
different activities have been emphasised, some of these differences
can be seen as clusters of performances with similar flux and entropy
in the two plots. For instance, the earliest series of rehearsals and
performances had flux between 0.05 and 0.15, while the period of
activity after July 2014 had flux between 0.1 and 0.25. This change
corresponded with an intensive series of study sessions and rehearsals
where the participants developed a more fluctuating style of
performance. The entropy between these two periods, however, remains
at a similiar level, showing the the space of available gestures that
the performers explored was more constant.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/flux-entropy-through-time}
  \caption{Distribution of flux and entropy values of sessions through time.
    Different performance styles are visible as clusters of similar
    flux regardless of performance context.
    \label{fig:flux-entropy-through-time}}
\end{figure}
